---
title: "The swedish housing market"
author: "Willie Langenberg"
date: '2021-02-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(plm)
library(corrplot)
```

## Abstract




## Introduction

The housing market vary a lot across different parts of a country. The houses being sold does not only differ size and design but also in pricing. I want to analyse and try to explain the general pricing of houses being sold in different counties in Sweden. With data containing information about the salary and population density in a given country, I will try to build a model for explaining the price difference of houses being sold in Sweden. 



## Data
```{r, echo=FALSE}
# read in and clean house price data
house_df <- read_csv("data/houseprices.csv",  locale = locale(encoding = "latin1"), col_types = cols())
house_df <- house_df %>%
  pivot_longer(-(1:2), names_to="year", values_to="y") %>%
  select(y, region, year)

# read in and clean salary data
salary_df <- read_csv("data/salary.csv", locale = locale(encoding = "latin1"), skip=2, col_types = cols())
salary_df <- salary_df %>%
  pivot_longer(-(1:4), names_to="year", values_to="salary") %>%
  select(salary, region, year)

# read in and clean population density data
popdensity_df <- read_csv("data/popdensity.csv", locale = locale(encoding = "latin1"), skip=2, col_types = cols()) 
popdensity_df <- popdensity_df %>%
  pivot_longer(-(1:2), names_to="year", values_to="pop_density") %>%
  select(pop_density, region, year)

# read in and clean population data
pop_df <- read_csv("data/population.csv", locale = locale(encoding = "latin1"), skip=2, col_types = cols()) 
pop_df <- pop_df %>%
  pivot_longer(-(1:2), names_to="year", values_to="population") %>%
  select(population, region, year)


#combine all data
df <- house_df %>% 
  inner_join(salary_df, by=c("year", "region")) %>%
  inner_join(popdensity_df, by=c("year", "region")) %>%
  inner_join(pop_df, by=c("year", "region"))

```

The data is downloaded from Statistics Sweden (scb.se). The response variable is the mean price of small houses being sold in a given year.  

## Analysis
```{r}
# Korrelationsplot 
corr_matrix <- cor(select(df, -region, -year))
corrplot(corr_matrix, type = "upper")
```


```{r}
# pairwise plots
df %>%
  mutate(y = 1000*y) %>%
  select(-year) %>%
  group_by(region) %>%
  pivot_longer(-(1:2), values_to = "x") %>%
  ggplot(aes(x = x, y = y, color=region)) +
  geom_point() + scale_y_log10() +
  facet_wrap(~name, scales = "free_x")
```


```{r}
# Ols regression
ols <- lm(y ~ salary + pop_density, data = df)
summary(ols)
```


```{r}
# fixed effects using least squares dummy variable model
fixed.dum <- lm(y ~ salary + pop_density + factor(region) - 1, data = df)
summary(fixed.dum)
```


```{r}
index <- sample(1:nrow(df), round(0.75 * nrow(df)))
train <- df[index, ]
test <- df[-index, ]

fixed_train <- lm(y ~ salary + pop_density + factor(region) - 1, data = train)
summary(fixed.dum)

pred.fix_lm <- predict(fixed_train, test)

### Plot Prediktion vs observerade värden
plot(pred.fix_lm, test$y, ylab="True values", xlab="Predicted values")
abline(0,1)
```


```{r}
# fixed effects
fixed <- plm(y ~ salary + pop_density, data = df, index=c("region", "year"), model="within" )
summary(fixed)

fixef(fixed)

pFtest(fixed, ols)
```


```{r}
# random effects
random <- plm(y ~ salary + pop_density, data = df, index=c("region", "year"), model="random" )
summary(random)


# Hausman test, fixed or random?
phtest(fixed, random)


# The p-value < 0.05 and we can therefore reject the null hypothesis that the random effects model is preferable over the fixed effects model.
```


# How should we now continue to analyze the data in a manner which ignores the time-dependence?



