---
title: "HW1"
author: "Willie Langenberg"
date: '2020-11-02'
output: github_document
---


## Past experiences
In past courses in mathematical statistics, we have always used R for writing the reports. With that said, I am somewhat familiar with the syntax in R, R Studio and R markdown. I have also used some "ggplot2", to plot nice looking graphics. However I have never used Git, Github or dplyr before.


## Data Example
Today the temperature felt unusually warm (14°C) for a november day in Stockholm. I figured I wanted to see some historical data for the temperature at this time of the year (2nd day of November). I simply used Google to search for "temperature dataset in sweden" and found "https://bolin.su.se/data/stockholm-thematic/air_temperature.php". It is a centre for climate research with open access datasets. On the website I chose to download data for the homogenized daily mean temperatures. The data came in the format of a textfile with tabseperated columns. To read it into R I used the function "read.table", which creates a dataframe to store the data in. The plot in _Figure.1_ below shows the historical mean temperature of every 2th of November day, from 1756 to 2005.
```{r, fig.cap="Plot that shows the mean temperature of each 2nd of November day, from 1756 to 2005."}
library(ggplot2)
df <-  read.table("daily_temp.txt")
df <- df[c(1, 2, 3, 6)]
names(df) <- c("Year", "Month", "Day", "Temperature")
new_df <- subset(df, df[2]==11 & df[3]==2)

ggplot(new_df, aes(Year, Temperature)) + geom_point()
```

_Figure 1:_ Plot that shows the mean temperature of each 2nd of November day, from 1756 to 2005.
```{r}
sessionInfo()
```


